Embedded Meta-Learning: Toward more flexible deep-learning models

### 2019-05-23 writing/embedded_meta_learning_main.tex
Polynomial example that would be nice to have.
Cut note
Cutting back to length

\section{Introduction}
Humans are able to use and reuse knowledge more flexibly than most deep learning models can \citep[e.g.][]{Lake2016, Marcus2018}. One fundamental reason for this is that humans are aware of what we are trying to compute and why. By contrast, there is a fundamental separation of knowledge within most deep neural networks -- although deep networks represent knowledge about data (in their activations) and knowledge about the structure of tasks (in their parameters), they do not represent any relationships between data and tasks. That is, a neural network's knowledge about what is being computed is only implicitly accessible to those computations. \par
There are a number of advantages to representing knowledge about data and tasks together. In particular, it can grant the ability to rapidly adapt behavior to a new task. The problem of rapid learning has been partially addressed by meta-learning systems \citep[e.g.][]{Vinyals2016, Finn2017a}. However, humans can use our knowledge of a task to flexibly alter our behavior in accordance with a change in task demands or a single instruction. For example, once we learn to play a game, we can immediately switch to playing in order to lose, and can achieve reasonable performance without any retraining (i.e. zero-shot). Deep learning systems at present generally lack this representational flexibility. \par
In this paper, we propose a new class of tasks based on this idea: meta-mappings, i.e. mappings between tasks (see below). This type of transfer is easily accessible to humans \citep{Lake2016}, but is generally inaccessible to most deep-learning models. To address this challenge, we propose a new class of architectures which essentially take a functional perspective on meta-learning, and exploit the idea of homoiconicity. (A homoiconic programming language is one in which programs in the language can be manipulated by programs in the language, just as data can.) By treating both data and task behaviors as functions, we can conceptually think of both data \emph{and} learned task behaviors as transformable. This yields the ability to not only learn to solve new tasks, but to learn how to transform these solutions in response to changing demands. We demonstrate that our architectures can flexibly remap their behavior to address the meta-mapping challenge. EML may also offer other benefits, such as a useful framework for continual learning. We suggest that approaches like EML will be key to building more intelligent and flexible deep learning systems. \par

\section{Meta-mapping}
%Can cut "the task"s in go sentence if needed
We propose the meta-mapping challenge. We define a meta-mapping as a task, or mapping, that takes a task as an input, output, or both. These include mappings from tasks to language (explaining), mappings from language to tasks (following instructions), and mappings from tasks to tasks (adapting behavior). While the first two categories have been partially addressed in prior work \citep[e.g.][]{Hermann2017, Co-Reyes2019}, the latter is more novel. \par
We argue that task-to-task meta-mappings are a useful way to think about human-like flexibility, because a great deal of our rapid adaptation is from a task to a variation on that task. For example, the task of playing go on a large board is closely related to the task of playing go on a small board. Humans can exploit this to immediately play well on a different board, but deep learning models generally have no way of achieving this. We can also adapt in much deeper ways, for example fundamentally altering our value function on a task, such as trying to lose. While meta-learning systems can rapidly learn a new task from a distribution of tasks they have experience with, this does not fully capture human flexibility. Given appropriate conditioning (see below), we can adapt to substantial task alterations zero-shot, that is, without seeing a single example from the new task \citep{Lake2016}. We suggest that meta-mappings offer a way to understand this flexibility, and a way for deep-learning models to achieve it.\par
There are different ways this adaptation can be cued, for example by examples of similar task mappings (``you've played checkers on a regular and half-size board, now try to adapt your go strategy to a smaller board,'') or simply by natural-language instructions (``play go on this smaller board'', see appendix \ref{app_why_meta_mapping} for a comparison to conditioning on a task description.). Achieving the flexibility to adapt to meta-map to new tasks will be an important step towards more general intelligence -- intelligence that is not limited to exactly the training domains it has seen. \par

\section{Embedded Meta-Learning (EML) architecture}
To address these challenges, we propose EML architectures, which are composed of two components: 
\vspace{-1em}
\begin{enumerate} \setlength \itemsep{0em}
\item An input/output system: domain specific encoders and decoders (vision, language, etc.) that map into a shared embedding space $Z$.
\item A meta-learning system which embeds tasks into the shared embedding space $Z$, and learns to map task embeddings to behavior.
\end{enumerate}
\vspace{-0.75em}
Why have a completely shared space $Z$ for data and tasks? The primary advantage is that it allows for arbitrary mappings between these entities. In addition to basic tasks, the system could in principle learn to perform meta-mappings to follow instructions or change behavior. This is a step closer to human-like flexibility. \par
Without training on meta-mappings, of course, the system will not be able to execute them well. However, if it is trained on a broad enough set of such mappings, it will be able to generalize to new instances drawn from the same data/task distribution, as we demonstrate below. For instances that fall outside its data distribution, or for optimal performance, it may require some retraining, however. This reflects the structure of human behavior -- we are able to learn rapidly when new knowledge is relatively consistent with our previous knowledge, but learning an entirely new paradigm (such as calculus for a new student) can be quite slow \citep[cf.][]{Kumaran2016}. \par 
\begin{figure}
\centering
\begin{tikzpicture}[auto, scale=0.75]
\node [dashblock] at (0, 0) (rep) {\begin{tabular}{c}$\bm{Z}$: \textbf{Shared} \\ \textbf{representation}\end{tabular}};

@@

@@

\paragraph{Basic meta-learning:} To perform a basic task, first, a training dataset is encoded by $\mathcal{I}$ and $\mathcal{T}$. These examples are fed to $\mathcal{M}$ to produce a function embedding. This function embedding is mapped through $\mathcal{H}$ to parameterize $F$, and then $F$ is used to process a (possibly distinct) dataset of embedded inputs, and $\mathcal{O}$ to map the resultant embeddings to outputs. This system can be trained end-to-end on targets for the second dataset. (Having two distinct datasets forces generalization at the meta-learning level, see appendix 
ref{app_clarifying_holdouts}.) See appendix \ref{app_model_details} for detailed architecture, operation, and hyper-parameters. \par 

@@

\paragraph{Meta-mapping:} However, the system is not limited to simply mapping inputs to outputs. The key advantage of our architecture is that it allows transforming anything that is embedded in $Z$ using the same system. Because tasks are embedded in $Z$, this allows for meta-mappings. For example, suppose we have an embedding $z_{game1} \in Z$ for the task of playing go. We can generate a meta-mapping embedding $z_{\text{meta}} \in Z$ from embeddings generated by the system when it is trying to win and lose various games: $z_{\text{meta}} = \mathcal{M}\left( \left\{\left((z_{game2},z_{game2,lose}\right), ... \right\}\right)$. We can generate a new task embedding $\hat{z}_{new} \in Z$:  

@@

This $\hat{z}_{new}$ could be interpreted as the system's guess at a losing strategy for game 1. We could then evaluate whether the system loses with this $\hat{z}_{new}$ strategy. All meta-mapping results reported here were evaluated in that way -- evaluating the loss on the target task after transforming the task embedding. \par

@@

%\subsection{Boolean functions}
%As a proof of concept, we first evaluated the system on a simple task of computing boolean functions on boolean inputs. Specifically, we can imagine mapping from binary-valued input vectors to a single binary output representing the evaluation of some logical proposition on the input. We train the system to figure out how to compute the predicate from seeing a subset of the (input, output) pairs. \par   
%We can then train the system to do various meta-mappings, for example identifying certain types of functions (like XOR) or negating predicates. \par
%
\section{Learning multivariate polynomials} \label{sec_poly}
\begin{figure}
\centering
\begin{subfigure}[t]{0.5\textwidth}
\includegraphics[width=\textwidth]{figures/poly/basic_results.png}
\caption{The polynomials domain, section \ref{sec_poly}.}
\label{poly_basic_results}
\end{subfigure}%
\begin{subfigure}[t]{0.5\textwidth}
\includegraphics[width=\textwidth]{figures/basic_meta_learning.png}
\caption{The cards domain, section \ref{sec_cards}.}
\label{cards_basic_results}
\end{subfigure}%
\caption{The EML system succeeds at the basic meta-learning tasks, which are a necessary prerequisite for meta-mappings. (\subref{poly_basic_results}) Basic meta-learning in the polynomials domain, section \ref{sec_poly}. The system successfully generalizes to held out polynomials. The solid line indicates optimal performance; the dashed line indicates chance performance (an untrained model). (\subref{cards_basic_results}) Basic meta-learning in the cards domain, section \ref{sec_cards}. The system successfully generalizes to held out games, both when trained on a random sample of half the tasks, or when a targeted subset is held out. The gray dashed line indicates chance performance, while the solid lines are optimal performance. The orange dashed lines show how well the system could perform by playing the strategy from the most correlated trained task. The fact that it generally exceeds this difficult baseline shows a deeper form of generalization than just memorizing a few strategies and picking the closest. Error-bars throughout are bootstrap 95\%-confidence intervals, numerical values for plots can be found in appendix \ref{app_numerical_results}.}
\end{figure}

@@

As a proof of concept, we first evaluated the system on the task of learning polynomials of degree $\leq 2$ in 4 variables (i.e. the task was to regress functions of the form $p: \mathbb{R}^4 \rightarrow \mathbb{R}$ where $p \in \mathcal{P}_2 \left(\mathbb{R}\right)$, though the model was given no prior inductive bias toward polynomial forms). This yields an infinite family of base-level tasks (the vector space of all such polynomials), as well as many families of meta-mappings over tasks (for example, multiplying polynomials by a constant, squaring them, or permuting their input variables). This allows us to not only examine the ability of the system to learn to learn polynomials from data, but also to adapt its learned representations in accordance with these meta-tasks. Details of the architecture and training can be found in appendix \ref{app_detailed_methods}.\par

@@

@@

@@

\newpage
% comment the following to compile without appendix but with references
\include{embedded_meta_learning_appendix}

>>>>

\vspace{-0.75em} % section
\section{Introduction}
\vspace{-0.5em} % section
Humans are able to use and reuse knowledge more flexibly than most deep learning models can \citep[e.g.][]{Lake2016, Marcus2018}. One fundamental reason for this is that humans are aware of what we are trying to compute and why. By contrast, there is a fundamental separation of knowledge within most deep neural networks -- although deep networks represent knowledge about data (in their activations) and knowledge about the structure of tasks (in their parameters), they do not represent any relationships between data and tasks. That is, a neural network's knowledge about what is being computed is only implicitly accessible to those computations. \par
There are a number of advantages to representing knowledge about data and tasks together. In particular, it can grant the ability to rapidly adapt behavior to a new task. The problem of rapid learning has been partially addressed by meta-learning systems \citep[e.g.][]{Vinyals2016, Finn2017a}. However, humans can use our knowledge of a task to flexibly alter our behavior in accordance with a change in task demands or a single instruction. For example, once we learn to play a game, we can immediately switch to playing in order to lose, and can achieve reasonable performance without any retraining (i.e. zero-shot). Deep learning systems at present generally lack this representational flexibility. \par
In this paper, we propose a new class of tasks based on this idea: meta-mappings, i.e. mappings between tasks (see below). This type of transfer is easily accessible to humans \citep{Lake2016}, but is generally inaccessible to most deep-learning models. To address this challenge, we propose a new class of architectures which essentially take a functional perspective on meta-learning, and exploit the idea of homoiconicity. (A homoiconic programming language is one in which programs in the language can be manipulated by programs in the language, just as data can.) By treating both data and task behaviors as functions, we can conceptually think of both data \emph{and} learned task behaviors as transformable. This yields the ability to not only learn to solve new tasks, but to learn how to transform these solutions in response to changing demands. We demonstrate that our architectures can flexibly remap their behavior to address the meta-mapping challenge. EML may also offer other benefits, such as a useful framework for continual learning. We suggest that approaches like EML will be key to building more intelligent and flexible deep learning systems. \par

\section{Meta-mapping}
\vspace{-0.5em} % section
%Can cut "the task"s in go sentence if needed
We propose the meta-mapping challenge. We define a meta-mapping as a task, or mapping, that takes a task as an input, output, or both. These include mappings from tasks to language (explaining), mappings from language to tasks (following instructions), and mappings from tasks to tasks (adapting behavior). While the first two categories have been partially addressed in prior work \citep[e.g.][]{Hermann2017, Co-Reyes2019}, the latter is more novel. \par
We argue that task-to-task meta-mappings are a useful way to think about human-like flexibility, because a great deal of our rapid adaptation is from a task to a variation on that task. For example, the task of playing go on a large board is closely related to the task of playing go on a small board. Humans can exploit this to immediately play well on a different board, but deep learning models generally have no way of achieving this. We can also adapt in much deeper ways, for example fundamentally altering our value function on a task, such as trying to lose. While meta-learning systems can rapidly learn a new task from a distribution of tasks they have experience with, this does not fully capture human flexibility. Given appropriate conditioning (see below), we can adapt to substantial task alterations zero-shot, that is, without seeing a single example from the new task \citep{Lake2016}. We suggest that meta-mappings offer a way to understand this flexibility, and a way for deep-learning models to achieve it.\par
There are different ways this adaptation can be cued, for example by examples of similar task mappings (``you've played checkers on a regular and half-size board, now try to adapt your go strategy to a smaller board,'') or simply by natural-language instructions (``play go on this smaller board'', see appendix \ref{app_why_meta_mapping} for a comparison to conditioning on a task description.). Achieving the flexibility to adapt to meta-map to new tasks will be an important step towards more general intelligence -- intelligence that is not limited to exactly the training domains it has seen. \par

\section{Embedded Meta-Learning (EML) architecture}
\vspace{-0.5em} % section
To address these challenges, we propose EML architectures, which are composed of two components: 
\vspace{-1em}
\begin{enumerate} \setlength \itemsep{0em}
\item An input/output system: domain specific encoders and decoders (vision, language, etc.) that map into a shared embedding space $Z$.
\item A meta-learning system which embeds tasks into the shared embedding space $Z$, and learns to map task embeddings to behavior.
\end{enumerate}
\vspace{-0.75em}
Why have a completely shared space $Z$ for data and tasks? The primary advantage is that it allows for arbitrary mappings between these entities. In addition to basic tasks, the system could in principle learn to perform meta-mappings to follow instructions or change behavior. \par
Without training on meta-mappings, of course, the system will not be able to execute them well. However, if it is trained on a broad enough set of such mappings, it will be able to generalize to new instances drawn from the same data/task distribution, as we demonstrate below. For instances that fall outside its data distribution, or for optimal performance, it may require some retraining, however. This reflects the structure of human behavior -- we are able to learn rapidly when new knowledge is relatively consistent with our previous knowledge, but learning an entirely new paradigm (such as calculus for a new student) can be quite slow \citep[cf.][]{Kumaran2016}. \par 
\begin{figure}
\centering
\begin{tikzpicture}[auto, scale=0.75]
\node [dashblock] at (0, 0) (rep) {\begin{tabular}{c}$\bm{Z}$: \textbf{Shared} \\ \textbf{representation}\end{tabular}};

@@

\looseness=-1

@@

\paragraph{Basic meta-learning:} To perform a basic task, first, a training dataset is encoded by $\mathcal{I}$ and $\mathcal{T}$. These examples are fed to $\mathcal{M}$ to produce a function embedding. This function embedding is mapped through $\mathcal{H}$ to parameterize $F$, and then $F$ is used to process a (possibly distinct) dataset of embedded inputs, and $\mathcal{O}$ to map the resultant embeddings to outputs. This system can be trained end-to-end on targets for the second dataset. (Having two distinct datasets forces generalization at the meta-learning level, see appendix \ref{app_clarifying_holdouts}.) See appendix \ref{app_model_details} for detailed architecture, operation, and hyper-parameters. \par 

@@

\paragraph{Meta-mapping:} However, the system is not limited to simply mapping inputs to outputs. Anything that is embedded in $Z$ can be transformed using the same system. Because tasks are embedded in $Z$, this allows for meta-mappings. For example, suppose we have an embedding $z_{game1} \in Z$ for the task of playing some game. We can generate a meta-mapping embedding $z_{\text{meta}} \in Z$ from embeddings generated by the system when it is trying to win and lose various games: $z_{\text{meta}} = \mathcal{M}\left( \left\{\left((z_{game2},z_{game2,lose}\right), ... \right\}\right)$. We can generate a new task embedding $\hat{z}_{new} \in Z$:  

@@

\looseness=-1
This $\hat{z}_{new}$ could be interpreted as the system's guess at a losing strategy for game 1. We could then evaluate whether the system loses with this $\hat{z}_{new}$ strategy. This is how we evaluate meta-mapping performance -- evaluating the loss on the target task after transforming the task embedding. \par

@@

%\subsection{Boolean functions}
%As a proof of concept, we first evaluated the system on a simple task of computing boolean functions on boolean inputs. Specifically, we can imagine mapping from binary-valued input vectors to a single binary output representing the evaluation of some logical proposition on the input. We train the system to figure out how to compute the predicate from seeing a subset of the (input, output) pairs. \par   
%We can then train the system to do various meta-mappings, for example identifying certain types of functions (like XOR) or negating predicates. \par
%
\section{Learning multivariate polynomials} \label{sec_poly}
\vspace{-0.5em} % section
\begin{figure}
\centering
\begin{subfigure}[t]{0.5\textwidth}
\includegraphics[width=\textwidth]{figures/poly/basic_results.png}
\caption{The polynomials domain, section \ref{sec_poly}.}
\label{poly_basic_results}
\end{subfigure}%
\begin{subfigure}[t]{0.5\textwidth}
\includegraphics[width=\textwidth]{figures/basic_meta_learning.png}
\caption{The cards domain, section \ref{sec_cards}.}
\label{cards_basic_results}
\end{subfigure}%
\caption{The EML system succeeds at the basic meta-learning tasks, which are a necessary prerequisite for meta-mappings. (\subref{poly_basic_results}) Basic meta-learning in the polynomials domain, section \ref{sec_poly}. The system successfully generalizes to held out polynomials. The solid line indicates optimal performance; the dashed line indicates chance performance (an untrained model). (\subref{cards_basic_results}) Basic meta-learning in the cards domain, section \ref{sec_cards}. The system successfully generalizes to held out games, both when trained on a random sample of half the tasks, or when a targeted subset is held out. The gray dashed line indicates chance performance, while the solid lines are optimal performance. The orange dashed lines show how well the system could perform by playing the strategy from the most correlated trained task. The fact that it generally exceeds this difficult baseline shows a deeper form of generalization than just memorizing a few strategies and picking the closest. Error-bars throughout are bootstrap 95\%-confidence intervals, numerical values for plots can be found in appendix \ref{app_numerical_results}.}
\end{figure}

@@

%% Cut example if necessary
As a proof of concept, we first evaluated the system on the task of learning polynomials of degree $\leq 2$ in 4 variables (i.e. the task was to regress functions of the form $p: \mathbb{R}^4 \rightarrow \mathbb{R}$ where $p \in \mathcal{P}_2 \left(\mathbb{R}\right)$, though the model was given no prior inductive bias toward polynomial forms). For example, if $p(w,x,y,z) = x$, the model might see examples like $(-1,1,1,1; 1)$ and $(0.7, 2, 1.3, -4; 2)$, and be evaluated on its output for points like $(-1, -1, 0.5, 0.3)$. This yields an infinite family of base-level tasks (the vector space of all such polynomials), as well as many families of meta-mappings over tasks (for example, multiplying polynomials by a constant, squaring them, or permuting their input variables). This allows us to not only examine the ability of the system to learn to learn polynomials from data, but also to adapt its learned representations in accordance with these meta-tasks. Details of the architecture and training can be found in appendix \ref{app_detailed_methods}.\par

@@

\vspace{-0.5em} % section

@@

%% Can cut The hand is explicitly ..., maybe some other stuff

@@

\newpage
% comment the following to compile without appendix but with references
%\include{embedded_meta_learning_appendix}
